---
title: "Practical Machine Learning Project"
author: "gah peng"
date: "December 26, 2015"
output: html_document
---

#Introduction

In this project, we will use the data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants who were asked to perform barbell lifts correctly and incorrectly in 5 different ways. The goal of this project is to predict the manner in which they did the exercise.

#Data
```{r, echo=FALSE}
rm(list = ls())
setwd("C:/gahpeng/coursera/08_machineLearning")

library(caret)
library(randomForest)
library(rpart)
library(rpart.plot)
```

##Data Loading

```{r}
trainDataRead <- read.csv(file = "pml-training.csv", na.strings=c("", "NA", "NULL"))
testDataRead <- read.csv(file = "pml-testing.csv", na.strings=c("", "NA", "NULL"))

```

##Removing empty or near empty columns

```{r}
trainDataReadCleaned <- trainDataRead[, colSums(is.na(trainDataRead))<15000]
trainDataReadCleaned <- trainDataReadCleaned[, colSums(trainDataReadCleaned=="")<15000]
```

##Removing not relevant columns

```{r}
trainDataReadCleaned<- trainDataReadCleaned[, -c(1, 2, 3, 4, 5, 6, 7)]
```

##Train and Test Data



```{r}
toTrainTestValidate <- createDataPartition(y=trainDataReadCleaned$classe, p=0.7, list=FALSE)
trainData <- trainDataReadCleaned[toTrainTestValidate,]
crossValidationData <- trainDataReadCleaned[-toTrainTestValidate,]
```

#Model Analysis
##(A) Random Forest

```{r, echo=FALSE}
set.seed(999)
modelRF <- randomForest(classe ~. , data=trainData, method="class")
myPrediction <- predict(modelRF, crossValidationData, type = "class")
confusionMatrix(myPrediction, crossValidationData$classe)
```


##(B) Decision Tree

```{r, echo=FALSE}
set.seed(999)
modelDTree <- rpart(classe ~ ., data=trainData, method="class")
prp(modelDTree,uniform=TRUE)
myPredictionDTree<- predict(modelDTree, crossValidationData, type = "class")
confusionMatrix(myPredictionDTree, crossValidationData$classe)

```


##Accuracy Rate

Comparing the accuracy rate for the two models, the accuracy rate for Random Forest is 99.54% while the accuracy rate for Decision Tree is 75.48%. 
Therefore the model selected is Random Forest.

##Out of Sample Error
Based on the missclassificaiton rate on the testing subset, an unbiased estimate of the random forest's out-of-sample error rate is 0.04%.


```{r, echo=FALSE}
missClass = function(values, prediction) {
  sum(prediction != values)/length(values)
}
errRate = missClass(crossValidationData$classe, myPrediction)
print(errRate)
```


## Variable of Importance

```{r, echo=FALSE} 
varImpPlot(modelRF)
```

#Test Model

```{r, echo=FALSE}
myTesting <- predict(modelRF, testDataRead)
print(myTesting)
```